### 2009-11-26 doc/manuals/python/SWIG-wrapped-ARC-Python-API.tex
Interchange DEBUG and VERBOSE loglevels.
Updated according to changes in code.

%%%%---- The title ----
{\centering \textsc{\Large The SWIG wrapped ARC Python API and the arcom utility package\footnote{as of rev. 13724}}\Large \par}
\vspace*{0.5cm}

@@

\subsubsection{Arc.i}
\label{iArc}
\begin{flushleft}
\begin{itemize}
  \item{\verb$%module arc$} \linebreak
    As stated in the SWIG documentation\footnote{\url{http://www.swig.org/Doc1.3/SWIGDocumentation.html}}: ''The \verb$%module$ directive defines the name of the module that will be created by SWIG.''
    That means, this module could be imported in Python code with: \textit{import arc}
\end{itemize}
\begin{itemize}
  \item{\verb$%include <stl.i>$} \linebreak
    Include the Standard Template Library, so that templates could be used.
\end{itemize}
\begin{itemize}
  \item{
\begin{verbatim}
  #ifdef SWIGPYTHON
  %include <std_list.i>
  #endif\end{verbatim}
    If target language is Python include \verb$std_list$.
  } \linebreak
\end{itemize}
\begin{itemize}
  \item{\verb$%template(StringList) std::list<std::string>;$\linebreak
    Define StringList as a template list of strings. An example of handling a list of strings is shown below. \linebreak
    \begin{example}
      \caption{Handling a list of strings}\label{listex}
\begin{verbatim}
>>> import arc
>>> #create an empty list
... sl = arc.StringList()
>>> #append strings to the list
... #size increases
... sl.size()
0
>>> sl.append('apple')
>>> sl.size()
1
>>> sl.append('banana')
>>> sl.size()
2
>>> sl.append('lemon')
>>> sl.size()
3
>>> sl.append('orange')
>>> sl.size()
4
>>> #list members
... sl[0], sl[1], sl[2], sl[3]
('apple', 'banana', 'lemon', 'orange')
>>> #do some slicing
... tmp = sl[:2]
>>> #now tmp contains the first two strings
... tmp.size()
2
>>> tmp[0], tmp[1]
('apple', 'banana')
>>> #do some more slicing
... tmp = sl[1:]
>>> #now tmp contains all the strings except the first
... tmp.size()
3
>>> tmp[0], tmp[1], tmp[2]
('banana', 'lemon', 'orange')
\end{verbatim}
    \end{example}
  }
\end{itemize}
\begin{itemize}
  \item{\verb$%template(StringStringMap) std::map<std::string, std::string>;$ \linebreak
    Define StringStringMap as a template map where both key and value are strings.\linebreak
    \begin{example}
      \caption{arc.StringStringMap}\label{mapex}
\begin{verbatim}
>>> import arc
>>> #create an empty map
... ssm = arc.StringStringMap()
>>> #add mapping
... ssm['key1'] = 'value1'
>>> ssm['key2'] = 'value2'
>>> #get keys
... ssm.keys()
['key1', 'key2']
>>> #get value for 'key1'
... ssm['key1']
'value1'
>>> #get value for 'key2'
... ssm['key2']
'value2'
\end{verbatim}
    \end{example}
  }
\end{itemize}
\begin{itemize}
  \item{
\begin{verbatim}
  %include "common.i"
  %include "message.i"
  %include "client.i"
  %include "data.i"
  %include "delegation.i"\end{verbatim}
  Include other interface files.
  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
  #ifdef SWIGPYTHON
  %include "security.i"
  #endif\end{verbatim}
  If target language is Python include interface file: security.i
  }
\end{itemize}
\end{flushleft}

%\newpage
\subsubsection{common.i}
\label{icommon}
\begin{flushleft}
\begin{itemize}
  \item{ \verb$%include <typemaps.i>$} \linebreak
  Typemaps will be used. 
\end{itemize}
\begin{itemize}
  \item{\begin{verbatim}
  %ignore operator !;
  %ignore operator [];
  %ignore operator =;
  %ignore operator ++;
  %ignore operator <<;

@@

@@

  Arc::URL objects. (See Example\ref{listex} for an example of using a list of objects - 

@@

\subsubsection{client.i}
\label{iclient}
\begin{flushleft}
\begin{itemize}
  \item{ \begin{verbatim}
  %template(ExecutionTargetList) std::list<Arc::ExecutionTarget>;
  %template(JobControllerList) std::list<Arc::JobController *>;
  %template(JobList) std::list<Arc::Job>;
  %template(JobStateList) std::list<Arc::JobState>;
  %template(SourceTypeList) std::list<Arc::DataSourceType>;
  %template(TargetTypeList) std::list<Arc::DataTargetType>;
  %template(FileTypeList) std::list<Arc::FileType>;
  %template(DirectoryTypeList) std::list<Arc::DirectoryType>;
  %template(ApplicationEnvironmentList) std::list<Arc::ApplicationEnvironment>;
  %template(SoftwareVersionList) std::list<Arc::SoftwareVersion>;
  %template(SoftwareRequirementList) std::list<Arc::SoftwareRequirement>;
  \end{verbatim}
  }
  Define ExecutionTargetList as a template list of Arc::ExecutionTarget objects, 
  JobControllerList as a template list of Arc::JobController pointers, 
  JobList as a template list of Arc::Job, JobStateList as a template list of 
  Arc::JobState, SourceTypeList as a template list of Arc::DataSourceType, 
  TargetTypeList as a template list of Arc::DataTargetType, FileTypeList as a 
  template list of Arc::FileType, DirectoryTypeList as a template list of 
  Arc::DirectoryType, ApplicationEnvironmentList as a template list of 
  Arc::ApplicationEnvironment, SoftwareVersionList as a template list of 
  Arc::SoftwareVersion and SoftwareRequirementList as a template list of 
  Arc::SoftwareRequirement objects. \linebreak
  (See Example\ref{listex} for an example of using a list of objects in Python.)
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
  #ifdef SWIGPYTHON
  namespace Arc {

  %typemap(in, numinputs=0) PayloadSOAP ** response (PayloadSOAP *temp) {
      $1 = &temp;
  }
  \end{verbatim}
  If target language is Python then this input typemap tells SWIG that ''response'' 
  arguments of type ''PayloadSOAP **'' found in ''Arc'' namespace are to be ignored. 
  As stated in the SWIG documentation: ''When numinputs is set to 0, the argument is 
  effectively ignored and cannot be supplied from the target language. '' 
  However the argument is still required when making the C/C++ call. This is solved 
  by providing a locally declared variable called temp from which the value used is 
  obtained. Statement ''\verb#$1 = &temp;#'' sets the input argument to point to this 
  temporary variable.
  \begin{verbatim}
  %typemap(argout) PayloadSOAP ** response {
      PyObject *o, *tuple;
      o = SWIG_NewPointerObj(*$1, SWIGTYPE_p_Arc__PayloadSOAP, SWIG_POINTER_OWN | 0 );
      tuple = PyTuple_New(2);
      PyTuple_SetItem(tuple,0,o);
      PyTuple_SetItem(tuple,1,$result);
      $result = tuple;
  }
  \end{verbatim}
  As the SWIG documentation tells: ''The \verb$"argout"$ typemap is used to return values from arguments.''
  With the previous typemap in mind this one indicates that the result will be changed as the following:
    \begin{itemize}
    \item{
      A tuple will be created with two members.
    }
    \item{
      The first member comes from the argument.
    }
    \item{
      The second one comes from the original result.
    }
    \item{
      The tuple will be returned as the new result.
    }
    \end{itemize}
  When processing a SOAP request it is required to return \verb$MCC_Status$ along with
  the SOAP response. This is achieved by returning a tuple that contains both of them.
  An example of calling the Echo service is shown below.
    \begin{example}
      \caption{ClientSOAP - process}\label{procex}
\begin{verbatim}
>>> import arc
>>> #create default config
... cfg = arc.MCCConfig()
>>> #create URL
... url = arc.URL('http://localhost:50000/Echo')
>>> #create payload
... payload = arc.PayloadSOAP(arc.NS({'echo':'urn:echo'}))
>>> #create payload content
... #and set echo message
... payload.NewChild('echo:echo').NewChild('echo:say').Set('Hello, World!')
>>> #create client
... client = arc.ClientSOAP(cfg,url)
>>> #let the client call the service
... response, status = client.process(payload)
>>> #see response
... print response.GetXML()
<soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.
org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" x
mlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XM
LSchema-instance"><soap-env:Body><echo:echoResponse><echo:hear>[ Hello, World! ]
</echo:hear></echo:echoResponse></soap-env:Body></soap-env:Envelope>
>>> #check status
... status.getKind() == arc.STATUS_OK
True
\end{verbatim}
    \end{example}

  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
%extend ACCLoader {
  Arc::Broker * getBroker(const std::string& brokername) { 
      return dynamic_cast<Arc::Broker*>( self->getACC(brokername) ); 
  }
  Arc::Broker * loadBroker(const std::string& brokername, const XMLNode * cfg) { 
      return dynamic_cast<Arc::Broker*>( self->loadACC(brokername, cfg) ); 
  }
  Arc::JobController * getJobController(const std::string& jobcontrollername) { 
      return dynamic_cast<Arc::JobController*>( self->getACC(jobcontrollername) ); 
  }
  Arc::JobController * loadJobController(const std::string& jobcontrollername, 
                                         const XMLNode * cfg) { 
      return dynamic_cast<Arc::JobController*>( 
                                         self->loadACC(jobcontrollername, cfg) ); 
  }
}
  \end{verbatim}
  SWIG methods added to get/load Broker and JobController objects from an ACCLoader object.
  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
  %apply std::string& INOUT { std::string& delegation_id };
  %include "../src/hed/libs/client/ClientX509Delegation.h"
  %clear std::string& delegation_id;
  \end{verbatim}
  This tells SWIG that all occurrences of ''\verb$delegation_id$'' of type ''\verb$std::string&$'' 
  as a parameter in ClientX509Delegation.h should be used as input and turned into 
  ''\verb$std::string&$'' output values as well. In fact, it depends on the type of delegation
  whether ''\verb$delegation_id$'' is used as an input or is created and returned as an output value.
  (See example below.)
    \begin{example}
      \caption{ClientX509Delegation - delegation\_id}\label{delegex}
\begin{verbatim}
For example the class ClientX509Delegation have a method createDelegation that makes 
use of delegation_id: 
  "For gridsite delegation service, the delegation_id
   is supposed to be created by client side, and sent to service side;
   for ARC delegation service, the delegation_id is supposed to be created
   by service side, and returned back. So for gridsite delegation service,
   this parameter is treated as input, while for ARC delegation service,
   it is treated as output."
However, this implies that one should take great care when dealing with delegation_id.
\end{verbatim}
    \end{example}
  }
\end{itemize}
\end{flushleft}

@@

  \begin{verbatim}
  %typemap(argout) std::list<FileInfo> & files {
    PyObject *o, *tuple;
    o = PyList_New(0);
    std::list<Arc::FileInfo>::iterator it;
    for (it = (*$1).begin(); it != (*$1).end(); ++it) {
        PyList_Append(o, SWIG_NewPointerObj(new Arc::FileInfo(*it), 
            SWIGTYPE_p_Arc__FileInfo, SWIG_POINTER_OWN | 0 ));
    }
    tuple = PyTuple_New(2);
    PyTuple_SetItem(tuple,0,o);
    PyTuple_SetItem(tuple,1,$result);
    $result = tuple;
  }
  \end{verbatim}
  This argout typemap - with the previous typemap in mind - indicates that the result will be changed as the following:
    \begin{itemize}
    \item{
      A tuple will be created with two members.
    }
    \item{
      The first member comes from the argument.
    }
    \item{
      The second one comes from the original result.
    }
    \item{
      The tuple will be returned as the new result.
    }
    \end{itemize}
    \begin{example}
      \caption{DataPointARC - ListFiles}\label{lfex}
\begin{verbatim}
One possible use of the FileInfo list could be the ListFiles funcionality of the 
DataPointARC class. This returns DataStatus along with the collected files. An example 
of using ListFiles will be shown later.
\end{verbatim} See Example\ref{url2dp} for details.
    \end{example}
  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
  %typemap(in, numinputs=0) std::string& failure_description (std::string temp) {
    $1 = &temp;
  }
  \end{verbatim}
  This input typemap tells SWIG that \verb$failure_description$ arguments of type \verb$std::string&$
  are to be ignored. However the argument is still required when making the C/C++ call. 
  This is solved by providing a locally declared variable called temp from which the value 
  used is obtained from. Statement ''\verb#$1 = &temp;#'' sets the input argument to point to this temporary variable.
  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
  %typemap(argout) std::string & failure_description {
    PyObject *o, *tuple;
    tuple = PyTuple_New(2);
    o = PyString_FromString($1->c_str());
    PyTuple_SetItem(tuple,0,$result);
    PyTuple_SetItem(tuple,1,o);
    $result = tuple;
  }
  \end{verbatim}
  This argout typemap - with the previous typemap in mind - indicates that the result will be changed as the following:
    \begin{itemize}
    \item{
      A tuple will be created with two members.
    }
    \item{
      The first member comes from the original result.
    }
    \item{
      The second one comes from the argument.
    }
    \item{
      The tuple will be returned as the new result.
    }

  One possible use of \verb$failure_description$ could be the Transfer funcionality of the 
  DataMover class. This returns a failure description along with the corresponding 
  DataStatus. This description may contain more details about the error if transfer 
  fails for some reason.
  (See example below.)
    \end{itemize}
    \begin{example}
      \caption{DataMover - Transfer}\label{dmex}
\begin{verbatim}
>>> import arc
>>> # create DataMover
... mover = arc.DataMover()
>>> # import datapoint_from_url from arcom package
... # DataPoints are required to use DataMover
... from arcom import datapoint_from_url
>>> # create source DataPoint
... src = datapoint_from_url('file://./file1')
>>> # create destination DataPoint
... dst = datapoint_from_url('file://./file2')
>>> # do the transfer
... status, failure = mover.Transfer(src, dst, arc.FileCache(), arc.URLMap())
>>> # check status
... str(status)
'Operation completed successfully'
>>> # check for failure message
... failure
''
\end{verbatim}
    \end{example}
  }
\end{itemize}
\begin{itemize}
  \item{\verb$%ignore Arc::DataHandle::operator->;$} \linebreak
  Ignored operator will not be accessible from Python.
\end{itemize}
\end{flushleft}
\subsubsection{delegation.i}
\label{idelegation}
\begin{flushleft}
\begin{itemize}
  \item{ \begin{verbatim}
  #ifdef SWIGPYTHON
  %ignore Arc::DelegationConsumer::Acquire(std::string&, std::string&);
  %ignore Arc::DelegationConsumerSOAP::UpdateCredentials(std::string&, std::string&, 
    const SOAPEnvelope&, SOAPEnvelope&);
  %ignore Arc::DelegationConsumerSOAP::DelegatedToken(std::string&, std::string&, 
    const SOAPEnvelope&, SOAPEnvelope&);
  %ignore Arc::DelegationConsumerSOAP::DelegatedToken(std::string&, std::string&, 
    const XMLNode&);
  %ignore Arc::DelegationContainerSOAP::UpdateCredentials(std::string&, 
    std::string&, const SOAPEnvelope&, SOAPEnvelope&);
  %ignore Arc::DelegationContainerSOAP::DelegatedToken(std::string&, std::string&, 
    const SOAPEnvelope&, SOAPEnvelope&);
  %ignore Arc::DelegationContainerSOAP::DelegatedToken(std::string&, std::string&, 
    const XMLNode&);
  #endif
  \end{verbatim}
  If target language is Python the selected methods are to be ignored and thus 
  will not be accessible.
  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
    %apply std::string& OUTPUT { std::string& credentials };
    %apply std::string& OUTPUT { std::string& identity };
    %ignore DelegationContainerSOAP::UpdateCredentials(
          std::string &,const SOAPEnvelope &, SOAPEnvelope &);
    %include "../src/hed/libs/delegation/DelegationInterface.h"
    %clear std::string& identity;
    %clear std::string& credentials;
  \end{verbatim}
  These tell SWIG that all occurrences of ''credentials'' and ''identity'' of type 
  ''\verb$std::string&$'' in DelegationInterface.h should be turned into 
  ''\verb$std::string&$'' output values.
  Furthermore, ''DelegationContainerSOAP::UpdateCredentials'' - with the parameters above - 
  will not be available from Python.
  Note that this is required by SWIG to properly generate Python code because applying the 
  previous typemap would result in two UpdateCredentials method having exactly the same parameter list.
    \begin{example}
      \caption{DelegationConsumerSOAP - UpdateCredentials}\label{dcsucex}
\begin{verbatim}
UpdateCredentials functionality of DelegationConsumerSOAP contains both credentials
and identity according to the C++ API documentation:

@@

Occurences are turned into output values, that is, Python API will have UpdateCredentials 
take two parameters (both of them being of type SOAPEnvelope) and return a list.
\end{verbatim}
    \end{example}
  }
\end{itemize}
\end{flushleft}

@@

    # set name of provided method
    self.request_names = ['foo']

@@

>>>>

%%%%---- The title ----
{\centering \textsc{\Large The SWIG wrapped ARC Python API and the arcom utility package}\Large \par}
\vspace*{0.5cm}

@@

\subsubsection{Arc.i}
\label{iArc}
\begin{flushleft}
\begin{itemize}
  \item{\verb$%module arc$} \linebreak
    As stated in the SWIG documentation\footnote{\url{http://www.swig.org/Doc1.3/SWIGDocumentation.html}}: ''The \verb$%module$ directive defines the name of the module that will be created by SWIG.''
    That means, this module could be imported in Python code with: \textit{import arc}
\end{itemize}
\begin{itemize}
  \item{\verb$%include <stl.i>$} \linebreak
    Include the Standard Template Library, so that templates could be used.
\end{itemize}
\begin{itemize}
  \item{
\begin{verbatim}
  #ifdef SWIGPYTHON
  %include <std_list.i>
  #endif\end{verbatim}
    If target language is Python include \verb$std_list$.
  } \linebreak
\end{itemize}
\begin{itemize}
  \item{\verb$%template(StringList) std::list<std::string>;$\linebreak
    Define StringList as a template list of strings. An example of handling a list of strings is shown below. \linebreak
    \begin{example}
      \caption{Handling a list of strings}\label{listex}
\begin{verbatim}
>>> import arc
>>> #create an empty list
... sl = arc.StringList()
>>> #append strings to the list
... #size increases
... sl.size()
0
>>> sl.append('apple')
>>> sl.size()
1
>>> sl.append('banana')
>>> sl.size()
2
>>> sl.append('lemon')
>>> sl.size()
3
>>> sl.append('orange')
>>> sl.size()
4
>>> #list members
... sl[0], sl[1], sl[2], sl[3]
('apple', 'banana', 'lemon', 'orange')
>>> #do some slicing
... tmp = sl[:2]
>>> #now tmp contains the first two strings
... tmp.size()
2
>>> tmp[0], tmp[1]
('apple', 'banana')
>>> #do some more slicing
... tmp = sl[1:]
>>> #now tmp contains all the strings except the first
... tmp.size()
3
>>> tmp[0], tmp[1], tmp[2]
('banana', 'lemon', 'orange')
\end{verbatim}
    \end{example}
  }
\end{itemize}
\begin{itemize}
  \item{\verb$%template(StringStringMap) std::map<std::string, std::string>;$ \linebreak
    Define StringStringMap as a template map where both key and value are strings.\linebreak
    \begin{example}
      \caption{arc.StringStringMap}\label{mapex}
\begin{verbatim}
>>> import arc
>>> #create an empty map
... ssm = arc.StringStringMap()
>>> #add mapping
... ssm['key1'] = 'value1'
>>> ssm['key2'] = 'value2'
>>> #get keys
... ssm.keys()
['key1', 'key2']
>>> #get value for 'key1'
... ssm['key1']
'value1'
>>> #get value for 'key2'
... ssm['key2']
'value2'
\end{verbatim}
    \end{example}
  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
#ifdef SWIGPYTHON
namespace Arc {

%typemap(in, numinputs=0) std::string& content (std::string str) { 
  $1 = &str; 
}
\end{verbatim}
  If target language is Python then this input typemap tells SWIG that ''content'' 
  arguments of type ''\verb$std::string&$'' found in ''Arc'' namespace are to be ignored. 
  As stated in the SWIG documentation: ''When numinputs is set to 0, the argument is 
  effectively ignored and cannot be supplied from the target language. '' 
  However the argument is still required when making the C/C++ call. This is solved 
  by providing a locally declared variable called str from which the value used is 
  obtained. Statement ''\verb#$1 = &str;#'' sets the input argument to point to this 
  temporary variable.
\begin{verbatim}
%typemap(argout) std::string& content { 
  PyObject *tuple; 
  tuple = PyTuple_New(2); 
  PyTuple_SetItem(tuple,0,$result); 
  PyTuple_SetItem(tuple,1,Py_BuildValue("s",$1->c_str())); 
  $result = tuple; 
} 
\end{verbatim}
  As the SWIG documentation tells: ''The \verb$"argout"$ typemap is used to return values from arguments.''
  With the previous typemap in mind this one indicates that the result will be changed as follows:
    \begin{itemize}
    \item{
      A tuple will be created with two members.
    }
    \item{
      The first member comes from the original result.
    }
    \item{
      The second one comes from the argument.
    }
    \item{
      The tuple will be returned as the new result.
    }
    \end{itemize}
  }
\end{itemize}
\begin{itemize}
  \item{
\begin{verbatim}
  %include "common.i"
  %include "message.i"
  %include "client.i"
  %include "data.i"
  %include "delegation.i"\end{verbatim}
  Include other interface files.
  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
  #ifdef SWIGPYTHON
  %include "security.i"
  #endif\end{verbatim}
  If target language is Python include interface file: security.i
  }
\end{itemize}
\end{flushleft}

%\newpage
\subsubsection{common.i}
\label{icommon}
\begin{flushleft}
\begin{itemize}
  \item{ \verb$%include <typemaps.i>$} \linebreak
  Typemaps will be used. 
\end{itemize}
\begin{itemize}
  \item{ \verb$%include <std_vector.i>$} \linebreak
  The \verb$std_vector$ library will be used.
\end{itemize}
\begin{itemize}
  \item{\begin{verbatim}
  %ignore operator !;
  %ignore operator [];
  %ignore operator =;
  %ignore operator ++;
  %ignore operator <<;

@@

  %template(URLVector) std::vector<Arc::URL>;
  %template(URLListMap) std::map< std::string, std::list<Arc::URL> >;

@@

  Arc::URL objects. Define URLVector as a template vector of Arc::URL objects and URLListMap 
  as a template map where the key is a string and the value is a list of Arc::URL objects. 
  (See Example\ref{listex} for an example of using a list of objects - 

@@

\subsubsection{client.i}
\label{iclient}
\begin{flushleft}
\begin{itemize}
  \item{ \begin{verbatim}
  %template(ExecutionTargetList) std::list<Arc::ExecutionTarget>;
  %template(JobControllerList) std::list<Arc::JobController *>;
  %template(JobList) std::list<Arc::Job>;
  %template(JobStateList) std::list<Arc::JobState>;
  %template(SourceTypeList) std::list<Arc::DataSourceType>;
  %template(TargetTypeList) std::list<Arc::DataTargetType>;
  %template(FileTypeList) std::list<Arc::FileType>;
  %template(DirectoryTypeList) std::list<Arc::DirectoryType>;
  %template(ApplicationEnvironmentList) std::list<Arc::ApplicationEnvironment>;
  %template(SoftwareList) std::list<Arc::Software>;
  %template(SoftwareRequirementList) std::list<Arc::SoftwareRequirement>;
  %template(ResourceTargetTypeList) std::list<Arc::ResourceTargetType>;
  \end{verbatim}
  }
  Define ExecutionTargetList as a template list of Arc::ExecutionTarget objects, 
  JobControllerList as a template list of Arc::JobController pointers, 
  JobList as a template list of Arc::Job, JobStateList as a template list of 
  Arc::JobState, SourceTypeList as a template list of Arc::DataSourceType, 
  TargetTypeList as a template list of Arc::DataTargetType, FileTypeList as a 
  template list of Arc::FileType, DirectoryTypeList as a template list of 
  Arc::DirectoryType, ApplicationEnvironmentList as a template list of 
  Arc::ApplicationEnvironment, SoftwareList as a template list of 
  Arc::Software, SoftwareRequirementList as a template list of 
  Arc::SoftwareRequirement and ResourceTargetTypeList as a template list of 
  Arc::ResourceTargetType objects. \linebreak
  (See Example\ref{listex} for an example of using a list of objects in Python.)
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
  #ifdef SWIGPYTHON
  namespace Arc {

  %typemap(in, numinputs=0) PayloadSOAP ** response (PayloadSOAP *temp) {
      $1 = &temp;
  }
  \end{verbatim}
  If target language is Python then this input typemap tells SWIG that ''response'' 
  arguments of type ''PayloadSOAP **'' found in ''Arc'' namespace are to be ignored. 
  As stated in the SWIG documentation: ''When numinputs is set to 0, the argument is 
  effectively ignored and cannot be supplied from the target language. '' 
  However the argument is still required when making the C/C++ call. This is solved 
  by providing a locally declared variable called temp from which the value used is 
  obtained. Statement ''\verb#$1 = &temp;#'' sets the input argument to point to this 
  temporary variable.
  \begin{verbatim}
  %typemap(argout) PayloadSOAP ** response {
      PyObject *o, *tuple;
      o = SWIG_NewPointerObj(*$1, SWIGTYPE_p_Arc__PayloadSOAP, SWIG_POINTER_OWN | 0 );
      tuple = PyTuple_New(2);
      PyTuple_SetItem(tuple,0,o);
      PyTuple_SetItem(tuple,1,$result);
      $result = tuple;
  }
  \end{verbatim}
  This argout typemap - with the previous typemap in mind - indicates that the result will be changed as follows:
    \begin{itemize}
    \item{
      A tuple will be created with two members.
    }
    \item{
      The first member comes from the argument.
    }
    \item{
      The second one comes from the original result.
    }
    \item{
      The tuple will be returned as the new result.
    }
    \end{itemize}
  When processing a SOAP request it is required to return \verb$MCC_Status$ along with
  the SOAP response. This is achieved by returning a tuple that contains both of them.
  An example of calling the Echo service is shown below.
    \begin{example}
      \caption{ClientSOAP - process}\label{procex}
\begin{verbatim}
>>> import arc
>>> #create default config
... cfg = arc.MCCConfig()
>>> #create URL
... url = arc.URL('http://localhost:50000/Echo')
>>> #create payload
... payload = arc.PayloadSOAP(arc.NS({'echo':'urn:echo'}))
>>> #create payload content
... #and set echo message
... payload.NewChild('echo:echo').NewChild('echo:say').Set('Hello, World!')
>>> #create client
... client = arc.ClientSOAP(cfg,url)
>>> #let the client call the service
... response, status = client.process(payload)
>>> #see response
... print response.GetXML()
<soap-env:Envelope xmlns:echo="urn:echo" xmlns:soap-enc="http://schemas.xmlsoap.
org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" x
mlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XM
LSchema-instance"><soap-env:Body><echo:echoResponse><echo:hear>[ Hello, World! ]
</echo:hear></echo:echoResponse></soap-env:Body></soap-env:Envelope>
>>> #check status
... status.getKind() == arc.STATUS_OK
True
\end{verbatim}
    \end{example}

  }
\end{itemize}
% \begin{itemize}
%   \item{ \begin{verbatim}
% %extend ACCLoader {
%   Arc::Broker * getBroker(const std::string& brokername) { 
%       return dynamic_cast<Arc::Broker*>( self->getACC(brokername) ); 
%   }
%   Arc::Broker * loadBroker(const std::string& brokername, const XMLNode * cfg) { 
%       return dynamic_cast<Arc::Broker*>( self->loadACC(brokername, cfg) ); 
%   }
%   Arc::JobController * getJobController(const std::string& jobcontrollername) { 
%       return dynamic_cast<Arc::JobController*>( self->getACC(jobcontrollername) ); 
%   }
%   Arc::JobController * loadJobController(const std::string& jobcontrollername, 
%                                          const XMLNode * cfg) { 
%       return dynamic_cast<Arc::JobController*>( 
%                                          self->loadACC(jobcontrollername, cfg) ); 
%   }
% }
%   \end{verbatim}
%   SWIG methods added to get/load Broker and JobController objects from an ACCLoader object.
%   }
% \end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
  %apply std::string& INOUT { std::string& delegation_id };
  %include "../src/hed/libs/client/ClientX509Delegation.h"
  %clear std::string& delegation_id;
  \end{verbatim}
  This tells SWIG that all occurrences of ''\verb$delegation_id$'' of type ''\verb$std::string&$'' 
  as a parameter in ClientX509Delegation.h should be used as input and turned into 
  ''\verb$std::string&$'' output values as well. In fact, it depends on the type of delegation
  whether ''\verb$delegation_id$'' is used as an input or is created and returned as an output value.
  (See example below.)
    \begin{example}
      \caption{ClientX509Delegation - delegation\_id}\label{delegex}
\begin{verbatim}
For example the class ClientX509Delegation have a method createDelegation that makes 
use of delegation_id: 
  "For gridsite delegation service, the delegation_id
   is supposed to be created by client side, and sent to service side;
   for ARC delegation service, the delegation_id is supposed to be created
   by service side, and returned back. So for gridsite delegation service,
   this parameter is treated as input, while for ARC delegation service,
   it is treated as output."
However, this implies that one should take great care when dealing with delegation_id.
\end{verbatim}
    \end{example}
  }
\end{itemize}
\end{flushleft}

@@

  \begin{verbatim}
  %typemap(argout) std::list<FileInfo> & files {
    PyObject *o, *tuple;
    o = PyList_New(0);
    std::list<Arc::FileInfo>::iterator it;
    for (it = (*$1).begin(); it != (*$1).end(); ++it) {
        PyList_Append(o, SWIG_NewPointerObj(new Arc::FileInfo(*it), 
            SWIGTYPE_p_Arc__FileInfo, SWIG_POINTER_OWN | 0 ));
    }
    tuple = PyTuple_New(2);
    PyTuple_SetItem(tuple,0,o);
    PyTuple_SetItem(tuple,1,$result);
    $result = tuple;
  }
  \end{verbatim}
  This argout typemap - with the previous typemap in mind - indicates that the result will be changed as follows:
    \begin{itemize}
    \item{
      A tuple will be created with two members.
    }
    \item{
      The first member comes from the argument.
    }
    \item{
      The second one comes from the original result.
    }
    \item{
      The tuple will be returned as the new result.
    }
    \end{itemize}
    \begin{example}
      \caption{DataPointARC - ListFiles}\label{lfex}
\begin{verbatim}
One possible use of the FileInfo list could be the ListFiles funcionality of the 
DataPointARC class. This returns DataStatus along with the collected files. An example 
of using ListFiles will be shown later.
\end{verbatim} See Example\ref{url2dp} for details.
    \end{example}
  }
\end{itemize}
\begin{itemize}
  \item{\verb$%ignore Arc::DataHandle::operator->;$} \linebreak
  Ignored operator will not be accessible from Python.
\end{itemize}
\end{flushleft}
\subsubsection{delegation.i}
\label{idelegation}
\begin{flushleft}
\begin{itemize}
  \item{ \begin{verbatim}
  #ifdef SWIGPYTHON
  %ignore Arc::DelegationConsumer::Acquire(std::string&, std::string&);
  %ignore Arc::DelegationConsumerSOAP::UpdateCredentials(std::string&, std::string&, 
    const SOAPEnvelope&, SOAPEnvelope&);
  %ignore Arc::DelegationConsumerSOAP::DelegatedToken(std::string&, std::string&, 
    const SOAPEnvelope&, SOAPEnvelope&);
  %ignore Arc::DelegationConsumerSOAP::DelegatedToken(std::string&, std::string&, 
    const XMLNode&);
  %ignore Arc::DelegationContainerSOAP::UpdateCredentials(std::string&, 
    std::string&, const SOAPEnvelope&, SOAPEnvelope&);
  %ignore Arc::DelegationContainerSOAP::DelegatedToken(std::string&, std::string&, 
    const SOAPEnvelope&, SOAPEnvelope&);
  %ignore Arc::DelegationContainerSOAP::DelegatedToken(std::string&, std::string&, 
    const XMLNode&);
  #endif
  \end{verbatim}
  If target language is Python the selected methods are to be ignored and thus 
  will not be accessible.
  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
    %apply std::string& OUTPUT { std::string& credentials };
    %apply std::string& OUTPUT { std::string& identity };
    %ignore DelegationContainerSOAP::UpdateCredentials(
          std::string &,const SOAPEnvelope &, SOAPEnvelope &);
    %include "../src/hed/libs/delegation/DelegationInterface.h"
    %clear std::string& identity;
    %clear std::string& credentials;
  \end{verbatim}
  These tell SWIG that all occurrences of ''credentials'' and ''identity'' of type 
  ''\verb$std::string&$'' in DelegationInterface.h should be turned into 
  ''\verb$std::string&$'' output values.
  Furthermore, ''DelegationContainerSOAP::UpdateCredentials'' - with the parameters above - 
  will not be available from Python.
  Note that this is required by SWIG to properly generate Python code because applying the 
  previous typemap would result in two UpdateCredentials method having exactly the same parameter list.
    \begin{example}
      \caption{DelegationConsumerSOAP - UpdateCredentials}\label{dcsucex}
\begin{verbatim}
UpdateCredentials functionality of DelegationConsumerSOAP contains both credentials
and identity according to the C++ API documentation:

@@

Occurences are turned into output values, that is, Python API will have UpdateCredentials 
take two parameters (both of them being of type SOAPEnvelope) and return a list.
\end{verbatim}
    \end{example}
  }
\end{itemize}
\begin{itemize}
  \item{ \begin{verbatim}
    %template(ScalableTimeInt) Arc::ScalableTime<int>;
    %template(RangeInt) Arc::Range<int>;
    %template(RangeInt64) Arc::Range<int64_t>;
  \end{verbatim}
  }
  These are template instantiations for template classes \verb$Arc::ScalableTime<int>$, 
  \verb$Arc::Range<int>$ and \verb$Arc::Range<int64_t>$.
\end{itemize}
\end{flushleft}

@@

    # set name of provided method
    self.request_names = ['foo','status']

@@

    return out

  def status(self, inpayload):
    # get local info
    dummyinfo = self.GetLocalInformation()

    # get service status
    dummystatus = str(dummyinfo.Get('AdminDomain').Get('Services').Get('Service').\
    Get('Endpoint').Get('ServingState'))

    # create response
    out = self._new_soap_payload()
    response_node = out.NewChild('dmmy:statusResponse')    
    response_node.Set('Status: ' + dummystatus)
